---
title: "R Notebook"
output: html_notebook
---
```{r}
## try http:// if https:// URLs are not supported
#source("https://bioconductor.org/biocLite.R")
#biocLite()
## try http:// if https:// URLs are not supported
#source("https://bioconductor.org/biocLite.R")
#biocLite("DESeq2")
print("loaded 1070820")
```
```{r}
library(DESeq2)
library(ggplot2)
library(reshape2)
library(lattice)

base_dir = "/home/heineike/github/expression_broad_data/expression_data/20170817_klscpka/"
```


```{r}

```

```{r}
#Load raw count data for each species from experiment.

countDataList = list(KL = NULL,SC  = NULL)
metaDataList = list(KL = NULL,SC  = NULL)

for (exp_species in c("KL","SC")){
  print(paste("species is ", exp_species))
  cts <- paste(base_dir,"20170817_klscpka_counts_",exp_species,".txt",sep = "")
  countData = read.table(cts, row.names = 1, header = TRUE, sep = "\t")
  #selecting out a subset: 
  # countData = countData[,c("7001","7002")]
   
  #this metadata file comes from a python routine. 
  meta <- paste(base_dir,"20170817_klscpka_metadata_deseq_",exp_species,".csv",sep = "")
  metaData = read.csv(meta, row.names = 1)
  colnames(countData) = row.names(metaData)
  countDataList[[exp_species]]=countData
  metaDataList[[exp_species]]=metaData
  print(metaData)
  print(countData)
  
  #Turns all kl imported data into a DESeqDataSet
  dds_all = DESeqDataSetFromMatrix(countData = countData, colData = metaData, design = ~strain+NMPP1+media+time)
  print(dds_all)
  #If there are any rows with zero or one count, this command removes them, (in this dataset there are not)
  paste(nrow(dds_all[ rowSums(counts(dds_all)) <= 1, ]), "items removed from Kl count data because there were no counts")
  dds_all_filtered = dds_all[ rowSums(counts(dds_all)) > 1, ]
  
  #Finds rlog values and prints to file

  #uses all default options except blind = FALSE instead of true 
  #Note from documentation:  
  #If many of genes have large differences in counts due to the experimental design,it is important to set blind=FALSE for downstream analysis.
  
  rld_all <- rlog(dds_all_filtered, blind = FALSE)  #, fitType = 'local')
  rlog_all <- assay(rld_all)
  #head(rlog_all)
  #plot(rlog_all[,"7050"],rlog_all[,"7002"])
  
  write.csv(as.data.frame(rlog_all), file=paste(base_dir,"20170817_klscpka_rlog_",exp_species,".csv",sep = ""))
  
  
}


# klMetaData <- metaData[c("AS1_minus","AS2_minus", "WT_minus","WT_plus", "AS1_plus","AS2_plus"),]
# #Combines final two columns of metadata
# klMetaData$condition = paste(klMetaData$strain,klMetaData$NMPP1,sep="_")
# klMetaData = klMetaData[ , "condition", drop = FALSE]
# klMetaData$condition = as.factor(klMetaData$condition)  #Avoid error message since this starts as a character. 
```


```{r}
# #Normalize raw count data using median of ratios method from DESEQ and compare with rlog data
# 
# #with a pseudo count 
# #klCountData_psuedo <- klCountData + 1/2
# #KiR <- apply(klCountData_psuedo,1,function(x) (prod(x))^(1/ncol(klCountData_psuedo)))
# #Kij_KiR  <- apply(klCountData_psuedo,2,function(x) x/KiR)
# 
# #Without a pseudo count
# KiR <- apply(klCountData,1,function(x) (prod(x))^(1/ncol(klCountData)))
# Kij_KiR  <- apply(klCountData,2,function(x) x/KiR)
# 
# #Visualizing histogram of each gene's ratio before picking the median. 
# Kij_KiR_melted <- melt(as.data.frame(Kij_KiR))
# ggplot(Kij_KiR_melted,aes(x = value)) + 
#     facet_wrap(~variable) + 
#     geom_histogram(binwidth = 0.1) +
#     coord_cartesian(xlim = c(0, 6)) 
# #WT Plus has a tighter distribution of counts - and that means that the sum which I was normalizing by before doesn't correlate well with the median of ratios. 
# sj <- apply(Kij_KiR,2, function(x) median(x,na.rm = TRUE))
# klCountData_norm <- t(apply(klCountData,1, function(x) x/sj))
# # scatterplot matrix 
# splom(klCountData_norm, 
#   	main="KL data",
#   	pscales = 0)
# 
# par(mfrow=c(2,3))
# xlabels = c("AS1_minus","AS1_minus", "WT_minus")
# ylabels = c("AS2_minus","AS1_plus", "WT_plus")
# for (jj in c(1:length(xlabels))){
#     plot(klCountData_norm[,xlabels[jj]],klCountData_norm[,ylabels[jj]], xlab = "",  ylab = ylabels[jj])
# }
# for (jj in c(1:length(xlabels))){
#     plot(rlog_all[,xlabels[jj]],rlog_all[,ylabels[jj]],xlab = xlabels[jj], ylab = ylabels[jj])
# }
# 
# write.csv(as.data.frame(klCountData_norm), 
#           file="C:/Users/Ben/Documents/GitHub/expression_broad_data/expression_data/kl_PKA_as_20160824/norm_KLacdata.csv")
```

```{r}
#Keep just the samples with SDC t=40, +/- NMPP1

samples_SDC_t40 <- list(KL = c("7001","7002","7005","7006","7039","7040"),SC  = c("7009","7010","7019","7020","7041","7042"))

for (exp_species in c("KL","SC")){
  print(paste("species is", exp_species))
  cts <- paste(base_dir,"20170817_klscpka_counts_",exp_species,".txt",sep = "")
  countData = read.table(cts, row.names = 1, header = TRUE, sep = "\t")
  
  exp_index = samples_SDC_t40[[exp_species]]
  
  countData = countDataList[[exp_species]]
  metaData = metaDataList[[exp_species]]
  
  
  countData_plusminus <- countData[,samples_SDC_t40[[exp_species]]]
  metaData_plusminus <- metaData[samples_SDC_t40[[exp_species]], "NMPP1" ,drop = FALSE]
  
  countData_plusminus
  metaData_plusminus
  
  
  dds_plusminus = DESeqDataSetFromMatrix(countData = countData_plusminus, colData = metaData_plusminus, design = ~NMPP1)
  #dds_plusminus$condition <- factor(dds_plusminus$condition, levels = c("noNMPP1", "yesNMPP1"))
  dds_plusminus_out = DESeq(dds_plusminus)
  plotDispEsts(dds_plusminus_out)
  
  #should I be removing genes with no counts? 
  
  res = results(dds_plusminus_out,  alpha = 10e-100)
  print(res)
  plotMA(res, main = paste(exp_species,"SDC t40 AS- VS AS+"), ylim = c(-6,6))
  #Sort results and write as a .csv file
  resOrdered <- res[order(res$padj),]
  write.csv(as.data.frame(resOrdered), file=paste(base_dir,"20170817_klscpka_DESEQ_SDC_t40_AsMinvAsPlus_",exp_species,".csv",sep = ""))
}

```


```{r}
#Remove strain with reporters from analysis (7050)
countData_norep = countData[ , -which(names(countData) %in% c("7050"))]
metaData_norep = metaData[-which(rownames(metaData) %in% c("7050")), ] 

dds_all = DESeqDataSetFromMatrix(countData = countData_norep, colData = metaData_norep, design = ~strain+NMPP1+media+time)
#If there are any rows with zero or one count, this command removes them, (in this dataset there are not)
dds_all
paste(nrow(dds_all[ rowSums(counts(dds_all)) <= 1, ]), "items removed from Kl count data because there were no counts")
dds_all_filtered = dds_all[ rowSums(counts(dds_all)) > 1, ]


dds_postdeseq = DESeq(dds_all_filtered)
res = results(dds_postdeseq)
```


```{r}
#Reduce experiment to just the two AS_minus and AS_plus experiments and get DESEQ fold changes

klCountData_plusminus <- klCountData[,c("AS1_minus","AS2_minus", "AS1_plus","AS2_plus")]
klMetaData_plusminus <- klMetaData[c("AS1_minus","AS2_minus","AS1_plus","AS2_plus"), ,drop = FALSE]

dds_plusminus = DESeqDataSetFromMatrix(countData = klCountData_plusminus, colData = klMetaData_plusminus, design = ~condition)
dds_plusminus$condition <- factor(dds_plusminus$condition, levels = c("AS_minus", "AS_plus"))
dds_plusminus_out = DESeq(dds_plusminus)
plotDispEsts(dds_plusminus_out)
res = results(dds_plusminus_out,  alpha = 10e-100)
resMLE <- results(dds_plusminus_out, addMLE=TRUE,  alpha = 10e-100)
plotMA(res, main = "KL AS- VS AS+", ylim = c(-6,6))
par(mfrow=c(1,2))
plotMA(res, main = "shrunken LFC", ylim = c(-6,6))
plotMA(resMLE, MLE=TRUE, main="unshrunken LFC", ylim=c(-6,6))
#Sort results and write as a .csv file
resOrdered <- res[order(res$padj),]
write.csv(as.data.frame(resOrdered), 
          file="C:/Users/Ben/Documents/GitHub/expression_broad_data/expression_data/kl_PKA_as_20160824/DESEQ_KLac_ASmin_ASplus.csv")
```

```{r}
#DESeq LFC data for KLac WT_minus/AS_minus/WT_plus v.s. KLac AS_plus 
#treating everything except AS_plus as a minus

klCountData_WTvAS <- klCountData[,c("WT_minus","AS1_plus","AS2_plus")]
klMetaData_WTvAS <- klMetaData[c("WT_minus","AS1_plus","AS2_plus"), , drop = FALSE]

dds_WTvAS = DESeqDataSetFromMatrix(countData = klCountData_WTvAS, colData = klMetaData_WTvAS, design = ~condition)
dds_WTvAS$condition <- factor(dds_WTvAS$condition, levels = c("WT_minus", "AS_plus"))

paste(nrow(dds_WTvAS[ rowSums(counts(dds_WTvAS)) <= 1, ]), "items removed from Kl count data because there were no counts")
dds_WTvAS_filtered = dds_WTvAS[ rowSums(counts(dds_WTvAS)) > 1, ]

dds_WTvAS_out = DESeq(dds_WTvAS_filtered)
plotDispEsts(dds_WTvAS_out)
res_WTvAS = results(dds_WTvAS_out,  alpha = 10e-8)
resMLE_WTvAS <- results(dds_WTvAS_out, addMLE=TRUE,  alpha = 10e-8)
plotMA(res_WTvAS, main = "KL WT- VS AS+", ylim = c(-6,6))
par(mfrow=c(1,2))
plotMA(res_WTvAS, main = "shrunken LFC", ylim = c(-6,6))
plotMA(resMLE_WTvAS, MLE=TRUE, main="unshrunken LFC", ylim=c(-6,6))
#Sort results and write as a .csv file
resOrdered_WTvAS <- res_WTvAS[order(res_WTvAS$padj),]
write.csv(as.data.frame(resOrdered_WTvAS),
           file="C:/Users/Ben/Documents/GitHub/expression_broad_data/expression_data/kl_PKA_as_20160824/DESEQ_KLac_WTmin_ASplus.csv")


```

Strange that the ratio of medians normalized data doesn't really seem to have a normalized sum: 
> apply(klCountData_psuedo,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
13836185 15081192 18936631 13403258 
> apply(klCountData_norm,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
14825647 14864233 17170150 13906631

Removing NA instead of adding a pseudocount: 
> apply(klCountData,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
13833701 15078708 18934147 13400774 
> apply(klCountData_norm,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
14820140 14861583 17166059 13902784 

I wonder why they used that? 

```{r}
source("https://bioconductor.org/biocLite.R")
biocLite("DESeq2")
```




Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).