---
title: "R Notebook"
output: html_notebook
---

```{r}
library(DESeq2)
library(ggplot2)
library(reshape2)
library(lattice)
#Would be nice to automatically change pasted os separators

```


```{r}
#DESeq LFC data for SCer WT_minus v.s. SC AS_plus
#Note: Not enough repeats to warrant using LFC


scCts <- "C:/Users/Ben/Documents/GitHub/expression_broad_data/expression_data/SCer_NMPP1_RNA_Seq/SC_1NMPP1_From_Kieran.tsv"
scCountData = read.table(scCts, row.names = 1, header = TRUE, sep = " ")
colnames(scCountData)<- c("WT_minus", "AS_plus", "WT_plus")
 
meta <- "C:/Users/Ben/Documents/GitHub/expression_broad_data/expression_data/SCer_NMPP1_RNA_Seq/sc_NMPP1_metadata.csv"
scMetaData = read.csv(meta, row.names = 1)
#Combines final two columns of metadata
scMetaData$condition = paste(scMetaData$strain,scMetaData$NMPP1,sep="_")
scMetaData = scMetaData[ , "condition", drop = FALSE]
scMetaData$condition = as.factor(scMetaData$condition)  #Avoid error message since this starts as a character. 
# 
# 
scCountData_WTvAS <- scCountData[,c("WT_minus","AS_plus")]
scMetaData_WTvAS <- scMetaData[c("WT_minus","AS_plus"), , drop = FALSE]
# 
sc_dds_WTvAS = DESeqDataSetFromMatrix(countData = scCountData_WTvAS, colData = scMetaData_WTvAS, design = ~condition)
sc_dds_WTvAS$condition <- factor(sc_dds_WTvAS$condition, levels = c("WT_minus", "AS_plus"))
# 
paste(nrow(sc_dds_WTvAS[ rowSums(counts(sc_dds_WTvAS)) <= 1, ]), "items removed from Kl count data because there were no counts")
sc_dds_WTvAS_filtered = sc_dds_WTvAS[ rowSums(counts(sc_dds_WTvAS)) > 1, ]
 
sc_dds_WTvAS_out = DESeq(sc_dds_WTvAS_filtered)
plotDispEsts(sc_dds_WTvAS_out)
sc_res_WTvAS = results(sc_dds_WTvAS_out,  alpha = 10e-8)
sc_resMLE_WTvAS <- results(sc_dds_WTvAS_out, addMLE=TRUE,  alpha = 10e-8)
plotMA(sc_res_WTvAS, main = "KL WT- VS AS+", ylim = c(-6,6))
par(mfrow=c(1,2))
plotMA(sc_res_WTvAS, main = "shrunken LFC", ylim = c(-6,6))
plotMA(sc_resMLE_WTvAS, MLE=TRUE, main="unshrunken LFC", ylim=c(-6,6))
#Sort results and write as a .csv file
# resOrdered_WTvAS <- res_WTvAS[order(res_WTvAS$padj),]
# write.csv(as.data.frame(resOrdered_WTvAS),
#            file="C:/Users/Ben/Documents/GitHub/expression_broad_data/expression_data/kl_PKA_as_20160824/DESEQ_KLac_WTmin_ASplus.csv")
# 


```

dds[ rowSums(counts(dds)) > 1, ]

library(reshape2)
library(ggplot2)
d <- melt(diamonds[,-c(2:4)])
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable,scales = "free_x") + 
    geom_histogram()

ggplot(vegLengths, aes(lengthg)) + geom_histogram(alpha = 0.5, aes(y = ..density..), position = 'identity')

Multi factor design: 
design(ddsMF) <- formula(~ type + condition)
ddsMF <- DESeq(ddsMF)

Set blind = FALSE

?? What's the difference between rlog and rlogTransformation

Example suggests just doing the command DESEQ


Commands from Kieran: 
dds_all = DESeqDataSetFromMatrix(countData = raw_counts, colData = meta, ~1)
rld_all <- rlogTransformation(dds_all)
rlogm_all <- assay(rld_all_filtered)

meanSdPlot(rlogm)
deseq.R code,

Strange that the ratio of medians normalized data doesn't really seem to have a normalized sum: 
> apply(klCountData_psuedo,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
13836185 15081192 18936631 13403258 
> apply(klCountData_norm,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
14825647 14864233 17170150 13906631

Removing NA instead of adding a pseudocount: 
> apply(klCountData,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
13833701 15078708 18934147 13400774 
> apply(klCountData_norm,2,sum)
AS1_plus AS2_plus WT_minus  WT_plus 
14820140 14861583 17166059 13902784 

I wonder why they used that? 

```{r}
source("https://bioconductor.org/biocLite.R")
biocLite("DESeq2")
```

Maybe I need to add type? (single read or paired end?)



Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).